I implemented both \ac{gsa} and \ac{gabsa} and attempted to replicate parts of the \ac{gsa}\cite{GSA} and \ac{gabsa}\cite{GABSA} papers. Let us see what I can conclude from the results, starting with \ac{gsa}.

In \cref{tab:res} there are two comparisons between my and the authors implementation of \ac{gsa}. Function $f_1$ optimized to a better value, the comparison using $f_1$ in \cref{fig:f1} shows that my implementation also converged faster. The hump at the start with my implementation is not present in the authors work. Clearly my implementation can not be their \ac{gsa}. Looking at $f_2$ my work converges 6 orders of magnitude worse then it should. I can only conclude that I did not successfully replicate \ac{gsa}.

The \ac{gabsa} paper \cite{GABSA} did not specify the cooling function used for annealing and its plots have little to no detail. The latter due to the strange choice for a linear scale\footnote{this hides almost every detail of the plot}. Additionally instead of reporting an average and variance the paper only mentions the best value reached among many runs. The paper does not mention the number of tries to reach this value.

Since my \ac{gabsa} work is based on my \ac{gsa} implementation which I just concluded was faulty we expect to perform weakly. From the plots we see my implementation has a huge variance between runs. We see (\cref{tab:res}) that our results are better for 3 of the 2 problems tried. Given the way the results are presented\footnote{frankly speaking I am disappointed this paper got through peer review} in the \ac{gabsa} paper I can not compare to it. I do not know whether my \ac{gsa} work comes close to a correct \ac{gabsa} implementation.
%
To conclude, neither \ac{gsa} and \ac{gabsa} where correctly replicated here, \ac{gabsa} can never be replicated exactly without access to the original data or source.

Given the state of the \ac{gabsa} paper\cite{GABSA} it should be recreated and the results there presented verified. Future work should include building \ac{gabsa} on top of a verified \ac{gsa} implementation then trying multiple cooling functions. Alternatively time could be spend fixing the here presented \ac{gsa} implementation.
